---
title: "Data Visualization - Group Project _ Jean"
author: "Jean Zwerger"
date: "2023-11-15"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center")
```

```{r, load_libraries, include = FALSE}
library(tidyverse)
library(lubridate)
library(vroom)
library(janitor)
library(skimr)
library(vroom)
library(mice) 
library(VIM)
library(countrycode)
library(rvest)
library(knitr)
library(readxl)
```


> **Goal:** Rank a predefined set of cities across multiple topics to define which are the best to live in. --\>

> **Guidelines** everyone is assigned to a topic and creates 2-3 for each need datasets that capture all cities in scope (careful about different methodologies) Suggestion: can find research papers instead of raw data --\>




# Cost of Living

To assess cost of living across the different cities in scope, we will look at three dimensions. How much income, living expenses, and price fluctuations (captured by inflation) are expected in the cities of choice.




## Income: Salary

In this section, we will put the average MAM graduate salary into the context of the chosen cities.

I tried the below two approaches to scrap data from the pdfs directly. Unfortunately, initial results were insufficient, so I proceeded to copy paste the pdf data into an excel before tidying up the data in R.


### LBS MAM Wages 2022: Experiment 1 - Pdf Extraction with Tesseract OCR

```{r, tesseract}

# Load the libraries
library(tesseract)
library(magrittr)

# Set up the Tesseract OCR engine
engine <- tesseract("eng")

# Path to the image file
image_path <- "C:/Users/jeanz/Documents/Data Visualization/MAM 2022.png" # Replace with your actual image path

# Perform OCR on the image
text <- ocr(image_path, engine = engine)
print(text)

# ======
# RESULT
# ======
# [1] "Salary information by industry\n| eee | rc\nIndustry Mean Max = Min. Mean Max Min.) Mean Max Min\noverall £49529 EN6IES £25461 $60.796 $1ESOG2 $2999 $69042 $179,427 $56085\nSalary information by location\nLocation Mean Max = Min Mean Max = Min. = Mean.) Max Min\nOverall £48529 EN69ES £26,461 $50.796 SIESOS2 $29,991 $69062 $179,427 $36083\n‘Africa/Middle East. «£47361 £54000 £34790 $58,793 $69970 $46200 $7482 $138.64 $850\nAsia £52707 £16965 £26,461 $66,289 $135,062 $29951 $83,061 $173427 $46390\nEurope (excl.UK) £49,474 £65000 E1215 SEISIG $79950 $36552 $ER182 $1164 $36,083\nUK £46741 €100.000 £55000 $8552 S278 $55865 $SL74 $7767 $57,659\n"
# ======

```


### LBS MAM Wages 2022: Experiment 2 - PDF Extraction with Tabulizer Table Extractor

```{r, tabulizer}

# Load the pdftools and tabulizer packages
library(pdftools)
library(tabulizer)

# Specify the path to the PDF file
pdf_path <- "C:/Users/jeanz/Documents/Data Visualization/MAM_Employment Report_2022.pdf" # Replace with the actual file path

# Extract tables from page 6
tables <- extract_tables(pdf_path, pages = 6)

# Print the extracted tables
print(tables)


# ======
# RESULT
# ======
# [[1]]
#      [,1]         [,2]      [,3]      [,4]      [,5]      [,6]      [,7]      [,8]      [,9]
# [1,] "CONSULTING" "£48,112" "£65,000" "£33,000" "$62,091" "$85,000" "$40,074" "$70,000" "$128,750"
#      [,10]
# [1,] "$42,056"
# 
# [[2]]
#      [,1]      [,2]      [,3]       [,4]      [,5]      [,6]       [,7]      [,8]      [,9]
# [1,] "FINANCE" "£59,469" "£116,965" "£36,000" "$72,599" "$135,062" "$42,678" "$82,021" "$179,427"
#      [,10]
# [1,] "$44,789"
# 
# [[3]]
#      [,1]          [,2]      [,3]      [,4]      [,5]      [,6]      [,7]      [,8]      [,9]
# [1,] "DIVERSIFIED" "£33,252" "£40,000" "£26,441" "$37,538" "$46,757" "$29,991" "$44,299" "$49,069"
#      [,10]
# [1,] "$37,639"
# ======

```


### LBS MAM Wages 2022: Loading and Tidying Up Data from Excel

```{r, loaddata_salary}

# LBS MAM employment reports 2020, 2021, 2022


# Load raw data
df <- read_csv("C:/Users/jeanz/Documents/Data Visualization/Employment Report _ Salary Report.csv")
glimpse(df)

# Clean the Data
# Identify 'Other compensation' rows
other_comp <- df %>%
  filter(Breakdown == "Other compensation") %>% 
  mutate(Index = row_number())

# Remove these rows from the original dataframe
main <- df %>%
  filter(Breakdown != "Other compensation") %>% 
  mutate(Index = row_number())

# Combine both dataframes by indexing corresponding rows
df_combined <- left_join(main, other_comp, by = c("Index"))

# Merge Mean, Min, Max columns together between base wages and other compensations to get total wages
df_transformed <- df_combined %>%
  # Drop specified columns
  select(-Index, -ends_with(".y")) %>%
  # Rename columns
  rename(Program = Programs.x, Year = Year.x, Category = Category.x, Breakdown = Breakdown.x,
         GBPMean_Value = Mean...5.x, GBPMax_Value = Max...7.x, GBPMin_Value = Min...9.x,
         USDMean_Value = Mean...11.x, USDMax_Value = Max...13.x, USDMin_Value = Min...15.x,
         INTMean_Value = Mean...17.x, INTMax_Value = Max...19.x, INTMin_Value = Min...21.x,
         GBPMean_Currency = Currency...6.x, GBPMax_Currency = Currency...8.x, GBPMin_Currency = Currency...10.x,
         USDMean_Currency = Currency...12.x, USDMax_Currency = Currency...14.x, USDMin_Currency = Currency...16.x,
         INTMean_Currency = Currency...18.x, INTMax_Currency = Currency...20.x, INTMin_Currency = Currency...22.x) %>%
  # Sum corresponding values
  mutate(GBPMean_Value = GBPMean_Value + df_combined$Mean...5.y,
         GBPMax_Value = GBPMax_Value + df_combined$Max...7.y,
         GBPMin_Value = GBPMin_Value + df_combined$Min...9.y,
         USDMean_Value = USDMean_Value + df_combined$Mean...11.y,
         USDMax_Value = USDMax_Value + df_combined$Max...13.y,
         USDMin_Value = USDMin_Value + df_combined$Min...15.y,
         INTMean_Value = INTMean_Value + df_combined$Mean...17.y,
         INTMax_Value = INTMax_Value + df_combined$Max...19.y,
         INTMin_Value = INTMin_Value + df_combined$Min...21.y)

# Pivot the "_Value" columns
df_values <- df_transformed %>%
  select(Program, Year, Category, Breakdown, ends_with("_Value")) %>%
  pivot_longer(
    cols = ends_with("_Value"),
    names_to = c("Currency", "Type"),
    names_pattern = "(GBP|USD|INT)(Mean|Max|Min)_Value",
    values_to = "Total_Wages",
    values_drop_na = TRUE
  )

# Pivot the "_Currency" columns
df_currency <- df_transformed %>%
  select(Program, Year, Category, Breakdown, ends_with("_Currency")) %>%
  pivot_longer(
    cols = ends_with("_Currency"),
    names_to = c("Currency", "Type"),
    names_pattern = "(GBP|USD|INT)(Mean|Max|Min)_Currency",
    values_to = "Currency_Type",
    values_drop_na = TRUE
  )

# Join the two dataframes based on their common columns
df_long <- left_join(df_values, df_currency, by = c("Program", "Year", "Category", "Breakdown", "Type", "Currency"))

# Convert Currency prefix to actual currency name
df_long <- df_long %>%
  mutate(Currency = case_when(
    Currency == "GBP" ~ "UK Pounds",
    Currency == "USD" ~ "American Dollars",
    Currency == "INT" ~ "International Dollars",
    TRUE ~ NA_character_
  )) %>%
  select(-Currency_Type) # Remove the redundant Currency_Type column

glimpse(df_long)


# Prepare data for visualization
# Filter out only the values needed for the visualization
Location_mean_total_wages_USD_2022 <- df_long %>% 
  filter(Currency == "American Dollars" & Type == "Mean" & Category == "Location" & Year == "2022")

# Define the mapping from 'Breakdown' to cities
breakdown_to_city <- c(
  "Overall" = "New York, United States of America; San Francisco, United States of America",
  "UK" = "London, United Kingdom",
  "Europe (excl. UK)" = "Berlin, Germany; Paris, France; Copenhagen, Denmark",
  "Africa/Middle East" = "Dubai, United Arab Emirates",
  "Asia" = "Singapore, Singapore; Shanghai, China; Tokyo, Japan; Mumbai, India; Sydney, Australia"
)

# Transform the dataframe
Location_mean_total_wages_USD_2022_cities <- df_long %>% 
  filter(Currency == "American Dollars" & Type == "Mean" & Category == "Location" & Year == 2022) %>%
  mutate(City = breakdown_to_city[Breakdown]) %>%
  separate_rows(City, sep = "; ") %>% # Split the cities into separate rows if multiple cities are
  rename('Mean Total Wage' = Total_Wages, Region = Breakdown) %>%
  separate(City, into = c("City", "Country"), sep = ", ", extra = "merge") %>%
  select(City, Country, Region, Currency, 'Mean Total Wage') %>% 
  arrange(desc(`Mean Total Wage`))

# Print "MAM Graduate Mean Wages across Geographies in 2022" table
print(Location_mean_total_wages_USD_2022_cities)


glimpse(Location_mean_total_wages_USD_2022_cities)

unique(Location_mean_total_wages_USD_2022_cities$City)

```


### Average Salary by City: Scrap and Load

```{r, }

# Alternative data source: OECD
# "Average salary by country"
# https://data.oecd.org/earnwage/average-wages.htm
# HOWEVER = not enough data in OECD source

# Source: https://www.numbeo.com/cost-of-living/prices_by_city.jsp?displayCurrency=USD&itemId=105

url_numbeo_average_salary <- "https://www.numbeo.com/cost-of-living/prices_by_city.jsp?displayCurrency=USD&itemId=105"

# Read the HTML content from the page
url_numbeo_average_salary_html <- read_html(url_numbeo_average_salary)

# Find the table using the specific CSS selector
url_numbeo_average_salary_table_node <- html_nodes(url_numbeo_average_salary_html, "table#t2")

# Extract the table into a dataframe
url_numbeo_average_salary_data <- html_table(url_numbeo_average_salary_table_node, fill = TRUE)

# The table is usually in the first list element
url_numbeo_average_salary_df <- url_numbeo_average_salary_data[[1]]

numbeo_average_monthly_salary_filtered <- url_numbeo_average_salary_df %>% 
  select(-Rank) %>%
  separate(City, into = c("City", "Country"), sep = ", ", extra = "merge") %>%
  mutate(City = trimws(City), Country = trimws(Country)) %>%
  filter(City %in% c("New York", "San Francisco",
                     "London", "Berlin", "Paris", "Copenhagen",
                     "Dubai",
                     "Singapore", "Shanghai", "Tokyo", "Mumbai",
                     "Sydney"))%>%
  mutate(Country = ifelse(City %in% c("New York", "San Francisco"), "United States of America", Country)) %>%
  filter(!(City == "London" & Country == "Canada"))

numbeo_average_annual_salary_filtered <- numbeo_average_monthly_salary_filtered %>% 
  mutate(`Average Annual Net Salary (After Tax)` = `Average Monthly Net Salary (After Tax)` * 12) %>% 
  select(-"Average Monthly Net Salary (After Tax)") %>% 
  arrange(desc(`Average Annual Net Salary (After Tax)`))

numbeo_average_annual_salary_filtered
glimpse(numbeo_average_annual_salary_filtered)

```


### Plot Salary Visualization

```{r, plot_salary}

# Join the two data frames on both "City" and "Country" columns
combined_df <- left_join(numbeo_average_annual_salary_filtered, Location_mean_total_wages_USD_2022_cities, by = c("City", "Country"))

# Select specific columns to view
selected_combined_df <- combined_df %>% 
  select(City, Country, Region, Currency, `Average Annual Net Salary (After Tax)`, `Mean Total Wage`) %>% 
  rename(
    `Average Annual Net Salary` = `Average Annual Net Salary (After Tax)`,
    `MAM Graduate Salary` = `Mean Total Wage`)

# Get a glimpse of the selected columns in the combined data frame
glimpse(selected_combined_df)


# Assuming selected_combined_df is your combined data frame with the correct columns
difference_df <- selected_combined_df %>%
  mutate(Difference = `MAM Graduate Salary` - `Average Annual Net Salary`)

# Function to create and adjust plot
create_adjusted_plot <- function() {
  ggplot(difference_df, aes(x = reorder(City, Difference), y = `Average Annual Net Salary`)) +
    geom_col(fill = "#D7E8CD") +  # This creates the bars for average annual net salary
    geom_point(aes(y = `MAM Graduate Salary`), color = "#0c530c", size = 5) +  # This adds the points for MAM graduate salary
    labs(title = "Average MAM Graduate Wage Compared to Average Net Wage per City",
       x = "",
       y = "Average Wage in US Dollars") +
    theme_minimal() +
    theme(plot.title = element_text(size = 36, hjust = 0),
          axis.title = element_text(size = 30),
          axis.text = element_text(size = 30)) +
    coord_flip()
}

# Create the plot
plot1_salary <- create_adjusted_plot()


# Print the plot to match the desired aspect ratio
# (this will only work if your graphics device window is resized to match the aspect ratio)
print(plot1_salary)

# Save the plot with the specified dimensions
ggsave("plot1_salary.png", plot = plot1_salary, width = 20, height = 10)


# # If Ordering by order of average salaries
# # Create the chart
# plot <- ggplot(selected_combined_df, aes(x = reorder(City, `Average Annual Net Salary`), y = `Average Annual Net Salary`)) +
#   geom_col(fill = "grey") +  # This creates the bars for average annual net salary
#   geom_point(aes(y = `MAM Graduate Salary`), color = "#08a045", size = 3) +  # This adds the points for MAM graduate salary
#   labs(title = "Comparison of Average Annual Net Salary and MAM Graduate Wage",
#        x = "USD",
#        y = "City") +
#   theme_minimal() +
#   coord_flip() +  # Flip the axes so cities are on the y-axis
#   theme(axis.text.y = element_text(angle = 0))  # Ensure y-axis labels are horizontal
# 
# plot


  
```




## Living Expenses: Shopping Basket

Let's now discover how much of our income is expected to be used on average living expenses in each city.


### Cost of Living: Scrap and Load the Data

```{r, scrapdata_cost_of_living}

# **Source:** <https://www.numbeo.com/cost-of-living/prices_by_city.jsp?displayCurrency=USD&itemId=101&itemId=100&itemId=228&itemId=224&itemId=60&itemId=66&itemId=64&itemId=62&itemId=110&itemId=118&itemId=121&itemId=14&itemId=19&itemId=17&itemId=15&itemId=11&itemId=16&itemId=113&itemId=9&itemId=12&itemId=8&itemId=119&itemId=111&itemId=112&itemId=115&itemId=116&itemId=13&itemId=27&itemId=26&itemId=29&itemId=28&itemId=114&itemId=6&itemId=4&itemId=5&itemId=3&itemId=2&itemId=1&itemId=7&itemId=105&itemId=106&itemId=44&itemId=40&itemId=42&itemId=24&itemId=20&itemId=18&itemId=109&itemId=108&itemId=107&itemId=206&itemId=25&itemId=30&itemId=33&itemId=34>


# URL of the page to scrape
url_numbeo_cost_of_living <- "https://www.numbeo.com/cost-of-living/prices_by_city.jsp?displayCurrency=USD&itemId=101&itemId=100&itemId=228&itemId=224&itemId=60&itemId=66&itemId=64&itemId=62&itemId=110&itemId=118&itemId=121&itemId=14&itemId=19&itemId=17&itemId=15&itemId=11&itemId=16&itemId=113&itemId=9&itemId=12&itemId=8&itemId=119&itemId=111&itemId=112&itemId=115&itemId=116&itemId=13&itemId=27&itemId=26&itemId=29&itemId=28&itemId=114&itemId=6&itemId=4&itemId=5&itemId=3&itemId=2&itemId=1&itemId=7&itemId=105&itemId=106&itemId=44&itemId=40&itemId=42&itemId=24&itemId=20&itemId=18&itemId=109&itemId=108&itemId=107&itemId=206&itemId=25&itemId=30&itemId=33&itemId=34"

# Read the HTML content from the page
numbeo_cost_of_living_html <- read_html(url_numbeo_cost_of_living)

# Find the table using the specific CSS selector
cost_of_living_table_node <- html_nodes(numbeo_cost_of_living_html, "table#t2")

# Extract the table into a dataframe
cost_of_living_data <- html_table(cost_of_living_table_node, fill = TRUE)

# The table is usually in the first list element
cost_of_living_df <- cost_of_living_data[[1]]

# Get the names of the dataframe
col_names <- names(cost_of_living_df)

# Rename the duplicate columns
col_names[6] <- "Domestic Beer (0.5 liter draught), Restaurant"
col_names[7] <- "Imported Beer (0.33 liter bottle), Restaurant"
col_names[16] <- "Domestic Beer (0.5 liter draught), Supermarket"
col_names[17] <- "Imported Beer (0.33 liter bottle), Supermarket"
col_names[8] <- "Coke/Pepsi (0.33 liter bottle), Restaurant"
col_names[9] <- "Water (0.33 liter bottle), Restaurant"

# Assign the new names back to the dataframe
names(cost_of_living_df) <- col_names

# Now remove the 'Rank' column
cost_of_living_df <- cost_of_living_df %>% select(-Rank)

# Split the City column into City and Country
cost_of_living_df <- cost_of_living_df %>%
  separate(City, into = c("City", "Country"), sep = ", ")

# Remove any leading or trailing spaces in City and Country columns
cost_of_living_df <- cost_of_living_df %>%
  mutate(City = trimws(City), Country = trimws(Country))

# Filter for specific cities if needed
cost_of_living_df_filtered <- cost_of_living_df %>%
  filter(City %in% c("New York", "San Francisco",
                     "London", "Berlin", "Paris", "Copenhagen",
                     "Dubai",
                     "Singapore", "Shanghai", "Tokyo", "Mumbai",
                     "Sydney"))

# Replace state abbreviations for specific cities with "United States of America"
cost_of_living_df_filtered <- cost_of_living_df_filtered %>%
  mutate(Country = ifelse(City %in% c("New York", "San Francisco") & 
                          Country %in% c("NY", "CA"), 
                          "United States of America", Country))

# Pivot Longer with Column Numbers
cost_of_living_df_tidy <- cost_of_living_df_filtered %>%
  pivot_longer(
    cols = -c(City, Country), 
    names_to = "Variable", 
    values_to = "Value",
    names_pattern = "(.*)",
    names_repair = "universal"
  ) %>%
  mutate(ColumnNumber = as.numeric(gsub("...\\.(.*)", "\\1", rownames(.))))

# Categorize Using grepl
cost_of_living_df_tidy <- cost_of_living_df_tidy %>%
  mutate(Category = case_when(
    grepl("Restaurant|McMeal|Cappuccino", Variable) ~ "Food & Beverages",
    grepl("Beer|Bottle of Wine|Cigarettes", Variable) ~ "Alcohol & Tobacco",
    grepl("Milk|Bread|Eggs|Cheese|Chicken|Beef|1.5 liter bottle|Apples|Oranges|Potato|Lettuce|Rice|Tomato|Banana|Onion", Variable) ~ "Groceries",
    grepl("Apartment|Price per Square Meter|Basic|Mortgage", Variable) ~ "Housing & Utilities",
    grepl("Internet|Mobile Phone", Variable) ~ "Communication",
    grepl("Fitness Club|Tennis Court|Cinema", Variable) ~ "Leisure & Recreation",
    grepl("Jeans|Summer Dress|Nike Running Shoes|Leather Business Shoes", Variable) ~ "Clothing & Footwear",
    grepl("Taxi|Gasoline|Volkswagen Golf|Toyota Corolla Sedan|Transport|Monthly Pass", Variable) ~ "Transportation",
    grepl("Preschool|Primary School", Variable) ~ "Education",
    grepl("Salary", Variable) ~ "Salary",
    TRUE ~ "Other"
  ))

# Reorder columns and drop ColumnNumber
cost_of_living_df_tidy <- cost_of_living_df_tidy %>%
  select(City, Country, Category, Variable, Value) %>% 
  arrange(City)

# Create a mapping data frame for Number_of_Purchases_per_Month
purchases_per_month <- data.frame(
  Variable = c(
    "Bottle of Wine (Mid-Range)", 
    "Domestic Beer (0.5 liter draught), Supermarket", 
    "Imported Beer (0.33 liter bottle), Supermarket", 
    "Cigarettes 20 Pack (Marlboro)", 
    "1 Pair of Jeans (Levis 501 Or Similar)", 
    "1 Summer Dress in a Chain Store (Zara, H&M, ...)", 
    "1 Pair of Nike Running Shoes (Mid-Range)", 
    "1 Pair of Men Leather Business Shoes", 
    "Internet (60 Mbps or More, Unlimited Data, Cable/ADSL)", 
    "Mobile Phone Monthly Plan with Calls and 10GB+ Data", 
    "Preschool (or Kindergarten), Full Day, Private, Monthly for 1 Child", 
    "International Primary School, Yearly for 1 Child", 
    "Meal, Inexpensive Restaurant", 
    "Meal for 2 People, Mid-range Restaurant, Three-course", 
    "McMeal at McDonalds (or Equivalent Combo Meal)", 
    "Domestic Beer (0.5 liter draught), Restaurant", 
    "Imported Beer (0.33 liter bottle), Restaurant", 
    "Coke/Pepsi (0.33 liter bottle), Restaurant", 
    "Water (0.33 liter bottle), Restaurant", 
    "Cappuccino (regular)", 
    "Milk (regular), (1 liter)", 
    "Loaf of Fresh White Bread (500g)", 
    "Eggs (regular) (12)", 
    "Local Cheese (1kg)", 
    "Water (1.5 liter bottle)", 
    "Chicken Fillets (1kg)", 
    "Apples (1kg)", 
    "Oranges (1kg)", 
    "Potato (1kg)", 
    "Lettuce (1 head)", 
    "Rice (white), (1kg)", 
    "Tomato (1kg)", 
    "Banana (1kg)", 
    "Onion (1kg)", 
    "Beef Round (1kg) (or Equivalent Back Leg Red Meat)", 
    "Apartment (1 bedroom) in City Centre", 
    "Apartment (1 bedroom) Outside of Centre", 
    "Apartment (3 bedrooms) in City Centre", 
    "Apartment (3 bedrooms) Outside of Centre", 
    "Basic (Electricity, Heating, Cooling, Water, Garbage) for 85m2 Apartment", 
    "Price per Square Meter to Buy Apartment in City Centre", 
    "Price per Square Meter to Buy Apartment Outside of Centre", 
    "Mortgage Interest Rate in Percentages (%), Yearly, for 20 Years Fixed-Rate", 
    "Fitness Club, Monthly Fee for 1 Adult", 
    "Tennis Court Rent (1 Hour on Weekend)", 
    "Cinema, International Release, 1 Seat", 
    "Average Monthly Net Salary (After Tax)", 
    "One-way Ticket (Local Transport)", 
    "Monthly Pass (Regular Price)", 
    "Gasoline (1 liter)", 
    "Volkswagen Golf 1.4 90 KW Trendline (Or Equivalent New Car)", 
    "Toyota Corolla Sedan 1.6l 97kW Comfort (Or Equivalent New Car)", 
    "Taxi Start (Normal Tariff)", 
    "Taxi 1km (Normal Tariff)", 
    "Taxi 1hour Waiting (Normal Tariff)"
  ),
  Number_of_Purchases_per_Month = c(
    2, 2, 4, 6, 1, 1, 0.5, 0.4, 
    1, 1, 0, 0, 2, 4, 4, 2, 4, 
    4, 6, 12, 3, 3, 3, 0.8, 12, 
    1.8, 0.8, 0.8, 2, 2, 2, 1, 
    1, 0.4, 1.2, 1, 0, 0, 0, 
    0.5, 0, 0, 0, 1, 4, 1, 0, 
    0, 1, 40, 0, 0, 4, 
    10, 0
  )
)

# Join the mapping with the cost of living data
cost_of_living_df_tidy <- left_join(cost_of_living_df_tidy, purchases_per_month, by = "Variable")

# Make sure categorization was successful
# Extract unique pairs of Category and Variable
 variable_category_pairs <- cost_of_living_df_tidy %>%
   select(Category, Variable, Number_of_Purchases_per_Month) %>%
   distinct() %>% 
   arrange(Category)
print(variable_category_pairs)

# Calculate Monthly Expenses
cost_of_living_df_monthly_tidy <- cost_of_living_df_tidy %>%
  mutate(Monthly_Expenses = Value * Number_of_Purchases_per_Month) %>% 
  select(-Value, - Number_of_Purchases_per_Month)

# View the updated dataframe
print(cost_of_living_df_monthly_tidy)

cost_of_living_df_monthly_tidy %>% 
  filter(City == "Singapore")


```


### Plot Cost of Living Visualization

```{r, plot_living_expenses}

# First, Order the cities based on the total Monthly_Expenses in descending order
cost_of_living_df_monthly_tidy <- cost_of_living_df_monthly_tidy %>%
  filter(Category != "Salary") %>%
  group_by(City) %>%
  mutate(Total_Expenses = sum(Monthly_Expenses)) %>%
  ungroup() %>%
  arrange(desc(Total_Expenses)) %>%
  mutate(City = factor(City, levels = unique(City)))

# Calculate the total expenses per city for labeling
total_monthly_expenses <- cost_of_living_df_monthly_tidy %>%
  group_by(City) %>%
  summarise(Total_Expenses = sum(Monthly_Expenses))

# Plot bar charts with categories stacked on one another for each city
plot2_living_expenses <- ggplot(cost_of_living_df_monthly_tidy, aes(x = City, y = Monthly_Expenses)) +
  geom_bar(aes(fill = Category), stat = "identity", position = "stack") +
  geom_text(
    data = total_monthly_expenses, 
    aes(x = City, y = Total_Expenses, label = Total_Expenses), # Position text at the end of the bars
    hjust = 0, # Align text to the start of the text element (left after coord_flip)
    color = "#0c530c", 
    size = 10,
    nudge_y = 0  # No vertical nudge required
  ) +
  scale_fill_brewer(palette = "Greens") +
  scale_y_continuous(limits = c(0, 7000)) + # Extend y-axis with a small margin
  labs(title = "Total Monthly Living Expenses by City",
       x = "",
       y = "Monthly Expenses (in US Dollars)",
       fill = "Category") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 36),
    axis.title.x = element_text(size = 30),
    axis.title.y = element_text(size = 30),
    axis.text.x = element_text(angle = 0, size = 30, vjust = 0.5),
    axis.text.y = element_text(size = 30),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 24),
    legend.position = c(0.95, 0.95), # Position inside the plot area
    legend.justification = c("right", "top"), # Anchor point for legend position
    legend.box.just = "right", # Justify the legend box inside the plotting area
    legend.key.size = unit(3, 'lines'),
    legend.margin = margin(t = 0, r = 10, b = 0, l = 5) # Adjust margins around the legend
  ) +
  coord_flip()

# Display the plot
print(plot2_living_expenses)

# Save the plot with the specified dimensions
ggsave("plot2_living_expenses.png", plot = plot2_living_expenses, width = 20, height = 20)

```




## Price Fluctuations: Inflation

After discovering expected income and living expenses, we want to now explore the stability of the given prices. To do so, we will look at historical inflation.


### Load IMF Inflation Data

```{r, loaddata_inflation}

# Source: https://data.imf.org/?sk=388DFA60-1D26-4ADE-B505-A05A558D9A42&sId=1479329132316

# Load IMF data from locally stored extraction
annual_inflation <- read_excel("C:/Users/jeanz/Documents/Data Visualization/Inflation-data.xlsx", sheet = "hcpi_a")

# Tidy up the data
annual_inflation_tidy <- annual_inflation %>%
  pivot_longer(
    cols = `1970`:`2022`,
    names_to = "Year", 
    values_to = "Inflation_Rate"
  ) %>%
  rename(CountryCode = 'Country Code') %>% 
  select(CountryCode, Country, Year, Inflation_Rate)  # Keep only relevant columns

# View the transformed data
glimpse(annual_inflation_tidy)

library(purrr) # for the map function

# Define a mapping from country codes to cities
country_code_to_city <- list(
  USA = c("New York", "San Francisco"),
  GBR = "London",
  DEU = "Berlin",
  FRA = "Paris",
  DNK = "Copenhagen",
  ARE = "Dubai",
  SGP = "Singapore",
  CHN = c("Shanghai"),
  JPN = "Tokyo",
  IND = "Mumbai",
  AUS = "Sydney"
)

# Apply the mapping to the inflation data
annual_inflation_filtered <- annual_inflation_tidy %>%
  filter(Year >= "2012", Year <= "2022") %>%
  rowwise() %>%
  mutate(City = list(country_code_to_city[[CountryCode]])) %>%
  unnest(City) %>% 
  select(Year, City, Country, Inflation_Rate) %>% 
  mutate(Year = as.numeric(Year))

# Print the result
print(annual_inflation_filtered)

```


### Plot Inflation Visualization

```{r, plot_inflation}

# First, calculate the average inflation rate for each country over the period
avg_inflation_by_country <- annual_inflation_filtered %>%
  group_by(Country) %>%
  summarise(Avg_Inflation = mean(Inflation_Rate, na.rm = TRUE))

# Identify the countries with the min and max average inflation rates
min_inflation_country <- avg_inflation_by_country %>%
  filter(Avg_Inflation == min(Avg_Inflation)) %>%
  pull(Country)

max_inflation_country <- avg_inflation_by_country %>%
  filter(Avg_Inflation == max(Avg_Inflation)) %>%
  pull(Country)

# Categorize the data
filtered_data <- annual_inflation_filtered %>%
  mutate(Category = case_when(
    Country == min_inflation_country ~ "Min: Japan",
    Country == max_inflation_country ~ "Max: India",
    TRUE ~ "Other Countries"
  ))

# Prepare data for the ribbon (min and max for Other Countries)
ribbon_data <- filtered_data %>%
  filter(Category == "Other Countries") %>%
  group_by(Year) %>%
  summarise(ymin = min(Inflation_Rate, na.rm = TRUE),
            ymax = max(Inflation_Rate, na.rm = TRUE))

# Create the plot
plot3_inflation <- ggplot(filtered_data, aes(x = Year, y = Inflation_Rate, group = Category, color = Category)) +
  geom_ribbon(data = ribbon_data, 
              aes(ymin = ymin, ymax = ymax, x = Year, group = NULL), 
              fill = "#D7E8CD", alpha = 0.6, inherit.aes = FALSE) +  # Draw ribbon first
  geom_line(data = filter(filtered_data, Category != "Other Countries"), size = 2) +  # Draw lines on top of ribbon
  scale_color_manual(values = c("Min: Japan" = "#35CF37", 
                                "Max: India" = "#0c530c", 
                                "Other Countries" = "#D7E8CD")) +
  labs(title = "Inflation Rate Extremes and Range Over the Past Decade",
       x = "Year",
       y = "Inflation Rate (%)",
       color = "Category") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 36),
    axis.title.x = element_text(size = 30),
    axis.title.y = element_text(size = 30),
    axis.text.x = element_text(angle = 0, size = 30, vjust = 0.5),
    axis.text.y = element_text(size = 30),
    legend.position = "right",
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 24),
    legend.key.size = unit(1, 'lines'),
    legend.margin = margin(t = 0, r = 10, b = 0, l = 10)
  ) +
  scale_x_continuous(breaks = seq(min(filtered_data$Year), max(filtered_data$Year), by = 1)) +
  scale_y_continuous(breaks = seq(from = 2 * floor(min(filtered_data$Inflation_Rate) / 2), 
                                  to = 2 * ceiling(max(filtered_data$Inflation_Rate) / 2), 
                                  by = 2))  # Ensure breaks at even numbers +
  coord_flip()

# Display the plot
print(plot3_inflation)

# Save the inflation plot with the specified dimensions to match the salary plot
ggsave("plot3_inflation.png", plot = plot3_inflation, width = 20, height = 10)

# Colors
# 32620e
# 9dcd5a
# 35CF37
# 
# 00FF03
# 0c530c
# b6da9f
# D7E8CD



```


## Ranking

```{r, cost}

# Calculate the difference and add a rank based on the Average Annual Net Salary
ranked_salary_df <- selected_combined_df %>%
  arrange(desc(`Average Annual Net Salary`)) %>%
  mutate(Rank = row_number()) %>%
  select(Rank, City, `Average Annual Net Salary`)

# View the ranked data frame
print(ranked_salary_df)

```

```{r, income}

# Rank the cities by the highest difference
ranked_cost_df <- difference_df %>%
  arrange(desc(Difference)) %>%
  mutate(Rank = row_number()) %>% 
  select(Rank, City, `Average Annual Net Salary`, Difference)

# View the final ranked data frame
print(ranked_cost_df)

```

```{r, inflation}

# List of cities for which we want to find the corresponding country and inflation rank
cities <- c("Singapore", "Shanghai", "Tokyo", "Mumbai", "Sydney", "New York",
            "San Francisco", "Berlin", "Paris", "Copenhagen", "London", "Dubai")

# Country to city mapping based on the provided list
city_to_country <- c(Singapore = "Singapore", Shanghai = "China", Tokyo = "Japan",
                     Mumbai = "India", Sydney = "Australia", 'New York' = "United States",
                     "San Francisco" = "United States", Berlin = "Germany", Paris = "France",
                     Copenhagen = "Denmark", London = "United Kingdom", Dubai = "United Arab Emirates")

# Create a data frame from the city to country mapping
city_country_df <- data.frame(City = names(city_to_country), Country = city_to_country)

# Arrange the cities by the average inflation rate in ascending order
ranked_inflation_country <- avg_inflation_by_country %>%
  arrange(Avg_Inflation) %>%
  mutate(Rank = row_number()) %>%  # Rank the cities from lowest to highest inflation
  select(Rank, Country, Avg_Inflation)

# Join the city_country_df with the inflation_rank to get the rank and average inflation for each city
city_inflation_df <- merge(city_country_df, ranked_inflation_country, by = "Country")

# Select and order the columns as needed
city_inflation_df <- city_inflation_df[, c("Rank", "City", "Country", "Avg_Inflation")]

# Arrange the data frame based on the Rank
city_inflation_df <- city_inflation_df[order(city_inflation_df$Rank),]

# View the ranked data frame
print(city_inflation_df)


```

```{r}

# Now flip ranking order to descending order (10 to 1 instead of 1 t 10)




# Join the data frames on the 'City' column
combined_df <- full_join(ranked_salary_df, ranked_cost_df, by = "City") %>%
  full_join(city_inflation_df, by = "City")

# Rename the rank columns for clarity
combined_df <- combined_df %>%
  rename(
    Salary_Rank = Rank.x,  # Replace with the actual column name from 'ranked_salary_df'
    Cost_Rank = Rank.y,    # Replace with the actual column name from 'ranked_cost_df'
    Inflation_Rank = Rank  # Replace with the actual column name from 'city_inflation_df'
  )

# Calculate the combined rank as the average of the individual ranks
combined_df <- combined_df %>%
  rowwise() %>%
  mutate(
    Average_Rank = mean(c(Salary_Rank, Cost_Rank, Inflation_Rank), na.rm = TRUE)
  ) %>%
  ungroup()

# Rank the cities based on the average rank in ascending order
combined_df <- combined_df %>%
  arrange(Average_Rank) %>%
  mutate(Combined_Rank = row_number())

# Select and reorder columns as needed
final_combined_df <- combined_df %>%
  select(Combined_Rank, City, Salary_Rank, Cost_Rank, Inflation_Rank, Average_Rank)

# View the final combined and ranked data frame
print(final_combined_df)

```



```{r}

# Join the data frames on the 'City' column
combined_df <- full_join(ranked_cost_df, city_inflation_df, by = "City")

# Rename the rank columns for clarity
combined_df <- combined_df %>%
  rename(
    Cost_Rank = Rank.x,    # Replace with the actual column name from 'ranked_cost_df'
    Inflation_Rank = Rank.y  # Replace with the actual column name from 'city_inflation_df'
  )

# Calculate the weighted average of the ranks
combined_df <- combined_df %>%
  rowwise() %>%
  mutate(
    Weighted_Average_Rank = (Cost_Rank * 1 + Inflation_Rank * 0.5) / 2.5
  ) %>%
  ungroup()

# Rank the cities based on the weighted average rank in descending order
combined_df <- combined_df %>%
  arrange(desc(Weighted_Average_Rank)) %>%
  mutate(Combined_Rank = row_number())

# Select and reorder columns as needed
final_combined_df <- combined_df %>%
  select(Combined_Rank, City, Cost_Rank, Inflation_Rank, Weighted_Average_Rank)

# View the final combined and ranked data frame
print(final_combined_df)


```


```{r}

# View the final combined and ranked data frame
print(ranked_salary_df)
```


